---
title: "Basic Node Types"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Basic Node Types}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo=TRUE,
  message=TRUE,
  warning=TRUE,
  eval=FALSE
)
library(improveR)
```


# Tutorial: Basic Node Types

This tutorial introduces the fundamental node types in the improve platform: Files, Folders, Links, and External Links. You'll learn how to create, manage, and work with these basic building blocks of the improve repository.

## Environment Variables

This tutorial requires the following environment variables to be set. 
We recommend creating a `.Renviron` file in your project root with these variables:

```{r env-vars-display, eval=FALSE}
# Add these lines to your .Renviron file:
# Sys.setenv(IMPROVER_REPO_URL = "https://<url>:<repoPort>/repository")
# Sys.setenv(IMPROVER_STEP = "<valid-entityId>")  # Ideally the entityId of your tutorial folder
```

## Tutorial Configuration

```{r tutorial-config}
# Define the test folder path for this tutorial
# Suggested path structure: /Projects/tutorials/<userName>/tutorial1
TUTORIAL_FOLDER <- "/Projects/tutorials/myUserName/tutorial1-basic-nodes"  # Update with your username

# Check if environment variables are set
if (Sys.getenv("IMPROVER_REPO_URL") == "") {
  stop("IMPROVER_REPO_URL environment variable is not set. Please set it in your .Renviron file.")
}

if (Sys.getenv("IMPROVER_STEP") == "") {
  stop("IMPROVER_STEP environment variable is not set. Please set it in your .Renviron file.")
}

# Display current configuration
cat("Repository URL:", Sys.getenv("IMPROVER_REPO_URL"), "\n")
cat("Step ID:", Sys.getenv("IMPROVER_STEP"), "\n")
cat("Tutorial folder:", TUTORIAL_FOLDER, "\n")
```

## Connect to Repository

```{r connect}
# Connect to improve repository
improveConnect()

# Verify connection is valid
# This is useful if your session has been idle or if tokens have expired
checkConnect()

# Enable editing capabilities
setEditable(TRUE)
```

## Verify Tutorial Folder

```{r verify-folder}
# Check if tutorial folder exists
tutorialResource <- tryCatch({
  loadResource(TUTORIAL_FOLDER)
}, error = function(e) {
  NULL
})

if (is.null(tutorialResource)) {
  stop(paste("Tutorial folder", TUTORIAL_FOLDER, "does not exist in the repository.",
             "Please create it first or use a different folder."))
}

if (tutorialResource$nodeType != "Folder") {
  stop(paste(TUTORIAL_FOLDER, "exists but is not a folder. Please use a folder path."))
}

cat("Tutorial folder found:", tutorialResource$path, "\n")
cat("Resource ID:", tutorialResource$resourceId, "\n")
cat("Entity ID:", tutorialResource$entityId, "\n")
cat("\nTip: You can use this Entity ID as your IMPROVER_STEP for better path resolution.\n")

# Check if folder is empty
childResources <- loadChildResources(tutorialResource)$data[[1]]

if (!is.null(childResources) && nrow(childResources) > 0) {
  cat("\nWarning: Tutorial folder is not empty. It contains", nrow(childResources), "items:\n")
  print(childResources[, c("name", "nodeType")])
  
  # Ask user if they want to continue
  response <- readline(prompt = "Do you want to continue anyway? (yes/no): ")
  if (tolower(response) != "yes") {
    stop("Tutorial cancelled. Please use an empty folder or clear the existing folder.")
  }
} else {
  cat("Tutorial folder is empty and ready for use.\n")
}

# Store tutorial folder resource for later use
TUTORIAL_RESOURCE <- tutorialResource
```

## 1. Working with Folders

Folders are containers for organizing files and other resources in the improve repository.

```{r create-folders}
# Create a folder structure for our examples
cat("\n=== Creating Folder Structure ===\n")

# Create main folders
dataFolder <- createFolder(
  folderName = "data", 
  targetIdent = TUTORIAL_RESOURCE$entityId,
  comment = "Folder for data files"
)
cat("Created folder:", dataFolder$path, "\n")

scriptsFolder <- createFolder(
  folderName = "scripts", 
  targetIdent = TUTORIAL_RESOURCE$entityId,
  comment = "Folder for R scripts"
)
cat("Created folder:", scriptsFolder$path, "\n")

outputsFolder <- createFolder(
  folderName = "outputs", 
  targetIdent = TUTORIAL_RESOURCE$entityId,
  comment = "Folder for output files"
)
cat("Created folder:", outputsFolder$path, "\n")

# Create a subfolder
rawDataFolder <- createFolder(
  folderName = "raw", 
  targetIdent = dataFolder$entityId,
  comment = "Raw data files"
)
cat("Created subfolder:", rawDataFolder$path, "\n")

# List all folders in tutorial folder
cat("\n=== Current Folder Structure ===\n")
children <- loadChildResources(TUTORIAL_RESOURCE)$data[[1]]
print(children[children$nodeType == "Folder", c("name", "path", "entityId")])
```

## 2. Working with Files

Files are standard digital files stored in the repository with full version control.

```{r create-files}
cat("\n=== Creating Files ===\n")

# Create a sample data file
sampleData <- data.frame(
  id = 1:10,
  value = rnorm(10),
  category = sample(c("A", "B", "C"), 10, replace = TRUE)
)

# Save locally first
write.csv(sampleData, "sample_data.csv", row.names = FALSE)

# Upload to repository
dataFile <- createFile(
  targetIdent = rawDataFolder$entityId,
  fileName = "sample_data.csv",
  localPath = "sample_data.csv",
  comment = "Sample dataset for tutorial"
)
cat("Created file:", dataFile$path, "\n")
cat("Entity ID:", dataFile$entityId, "\n")
cat("Version ID:", dataFile$entityVersionId, "\n")

# Create an R script file
scriptContent <- '# Simple analysis script
data <- read.csv("sample_data.csv")
summary(data)
plot(data$value, main = "Sample Data Values")
'

writeLines(scriptContent, "analysis_script.R")

scriptFile <- createFile(
  targetIdent = scriptsFolder$entityId,
  fileName = "analysis_script.R",
  localPath = "analysis_script.R",
  comment = "Basic analysis script"
)
cat("\nCreated script:", scriptFile$path, "\n")

# Clean up local files
unlink(c("sample_data.csv", "analysis_script.R"))
```

## 3. File Versioning

Every change to a file creates a new version, preserving the complete history.

```{r file-versioning}
cat("\n=== File Versioning ===\n")

# Modify the data file
modifiedData <- sampleData
modifiedData$value <- modifiedData$value * 2
modifiedData$new_column <- runif(10)

write.csv(modifiedData, "sample_data_v2.csv", row.names = FALSE)

# Update the file (creates new version)
updatedFile <- updateFileContent(
  ident = dataFile$entityId,
  localPath = "sample_data_v2.csv",
  comment = "Updated data with new column and scaled values"
)

cat("Original version:", dataFile$entityVersionId, "\n")
cat("New version:", updatedFile$entityVersionId, "\n")
cat("Entity ID (unchanged):", updatedFile$entityId, "\n")

# View file history
history <- loadHistory(dataFile$entityId)$data[[1]]
cat("\n=== File History ===\n")
print(history[, c("entityVersionId", "lastModifiedOnDate", "comment")])

# Clean up
unlink("sample_data_v2.csv")
```

## 4. Working with Links

Links are stable pointers to specific versions of resources. They ensure referential integrity by always pointing to the same version unless explicitly updated.

```{r create-links}
cat("\n=== Creating Links ===\n")

# Create a link to the current version of our data file
dataLink <- createLink(
  links = updatedFile$entityVersionId,
  linkName = "stable_data_link",
  linkContainer = outputsFolder$resourceId
)
cat("Created link:", dataLink$path, "\n")
cat("Link points to version:", dataLink$targetRevisionId, "\n")
cat("Target file version:", updatedFile$entityVersionId, "\n")

# Now update the data file again
cat("\n=== Updating the File Again ===\n")
finalData <- modifiedData
finalData$value <- finalData$value * 1.5
finalData$final_column <- "final_version"

write.csv(finalData, "sample_data_v3.csv", row.names = FALSE)

# Create another new version
finalFile <- updateFileContent(
  ident = dataFile$entityId,
  localPath = "sample_data_v3.csv",
  comment = "Final update with additional changes"
)

cat("New file version created:", finalFile$entityVersionId, "\n")

# Check that the link still points to the old version
cat("\n=== Link Stability After File Update ===\n")
linkCheck <- updateResource(dataLink$entityId)
cat("Link still points to version:", linkCheck$targetRevisionId, "\n")
cat("Current file version:", finalFile$entityVersionId, "\n")
cat("Link is stable: ", linkCheck$targetVersionId == updatedFile$resourceVersionId, "\n")

# Update the link to point to the new version
cat("\n=== Updating the Link ===\n")
updatedLink <- updateLinks(
  links = dataLink$entityId,
  comment = "Updated link to point to latest version"
)
updatedLink <- updateResource(dataLink$entityId)
cat("Link now points to version:", updatedLink$resourceVersionId, "\n")
cat("Matches current file version:", updatedLink$targetVersionId == finalFile$resourceVersionId, "\n")

# Clean up
unlink("sample_data_v3.csv")
```

## 5. Working with External Links

External links point to resources outside the improve repository, specified by URLs.

```{r external-links}
cat("\n=== Creating External Links ===\n")

# Create an external link to documentation
docLink <- createExternalLink(
  url = "https://scinteco.com",
  linkName = "Scinteco Webpage",
  targetIdent =  TUTORIAL_RESOURCE$resourceId
)
cat("Created external link:", docLink$path, "\n")
cat("URL:", docLink$url, "\n")

# Create an external link to a data source
dataSourceLink <- createExternalLink(
  url = "https://data.gov/dataset/example",
  linkName = "External_Data_Source",
  targetIdent = dataFolder$resourceId
)
cat("\nCreated external link:", dataSourceLink$path, "\n")
```

::: {#fig-createExternalLink layout-ncol=1 .no-margin .left-caption}

![Improve Client: An external link is shows in the inventory folder of, e.g., a step. The entity (version) id features
the abbreviation 'EL'](articles/img/createExternalLink.png)

:::


## 6. Exploring Resource Properties

Every resource in improve has standard properties that help track and manage it.

```{r resource-properties}
cat("\n=== Resource Properties ===\n")

# Load a resource and examine its properties
resource <- loadResource(dataFile$entityId)

cat("Resource Name:", resource$name, "\n")
cat("Node Type:", resource$nodeType, "\n")
cat("Entity ID:", resource$entityId, "\n")
cat("Entity Version ID:", resource$entityVersionId, "\n")
cat("Resource ID:", resource$resourceId, "\n")
cat("Path:", resource$path, "\n")
cat("Created At:", as.character(resource$createdAtDate), "\n")
cat("Created By:", resource$createdByName, "\n")
cat("Is Version:", resource$isVersion, "\n")

# Check metadata
metadata <- loadMetaData(resource$entityId)$data[[1]]
if (!is.null(metadata) && nrow(metadata) > 0) {
  cat("\n=== Metadata ===\n")
  print(metadata)
}
```

## 7. Summary and Best Practices

```{r summary}
cat("\n=== Summary of Created Resources ===\n")

# List all resources created in this tutorial
allChildren <- loadChildResources(TUTORIAL_RESOURCE)$data[[1]]
resourceSummary <- allChildren[, c("name", "nodeType", "path")]
print(resourceSummary)

cat("\n=== Key Concepts Covered ===\n")
cat("1. Folders - Organize resources hierarchically\n")
cat("2. Files - Store data with automatic versioning\n")
cat("3. Links - Create stable references to specific versions\n")
cat("4. External Links - Reference resources outside improve\n")
cat("5. Versioning - Every change creates an immutable snapshot\n")
cat("6. Entity IDs - Unique identifiers for precise referencing\n")

cat("\n=== Best Practices ===\n")
cat("- Use descriptive names and comments for all resources\n")
cat("- Organize files in a logical folder structure\n")
cat("- Use links to ensure reproducibility by pointing to specific versions\n")
cat("- Track file history to understand how content has evolved\n")
cat("- Use external links for references that shouldn't be stored in improve\n")
```

## Clean Up

```{r cleanup}
# Note: We're leaving the created resources in place for educational purposes
# In a real scenario, you might want to clean up test resources

# Disconnect from improve
improveDisconnect()

cat("\nTutorial completed successfully!\n")
cat("All created resources remain in:", TUTORIAL_FOLDER, "\n")
cat("You can explore them further using the improve web interface.\n")
```
