---
title: "CLI tools"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CLI tools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo=TRUE,
  message=TRUE,
  warning=TRUE,
  eval=FALSE
)
library(improveR)
```


# CLI Tools

This tutorial demonstrates the CLI-style tools in improveR for working with local copies of repository content. You'll learn how to clone folders, push changes, pull updates, and manage resource locks.

**Prerequisites**: Complete Tutorial 01 first, as this tutorial expects the folder structure and files created there.

## Environment Variables

This tutorial requires the following environment variables to be set:

```{r env-vars-display, eval=FALSE}
# Add these lines to your .Renviron file:
# Sys.setenv(IMPROVER_REPO_URL = "https://<url>:<repoPort>/repository")
# Sys.setenv(IMPROVER_STEP = "<valid-entityId>")  # Ideally the entityId of your tutorial folder
```

## Tutorial Configuration

```{r tutorial-config}
# Define the paths for this tutorial
TUTORIAL_FOLDER <- "/Projects/tutorials/myUserName/tutorial1-basic-nodes"  # Update with your username
CLI_WORKSPACE <- "./tutorial2-workspace"  # Local directory for CLI operations

# Check if environment variables are set
if (Sys.getenv("IMPROVER_REPO_URL") == "") {
  stop("IMPROVER_REPO_URL environment variable is not set. Please set it in your .Renviron file.")
}

if (Sys.getenv("IMPROVER_STEP") == "") {
  stop("IMPROVER_STEP environment variable is not set. Please set it in your .Renviron file.")
}

# Display current configuration
cat("Repository URL:", Sys.getenv("IMPROVER_REPO_URL"), "\n")
cat("Step ID:", Sys.getenv("IMPROVER_STEP"), "\n")
cat("Tutorial folder:", TUTORIAL_FOLDER, "\n")
cat("Local workspace:", CLI_WORKSPACE, "\n")

# Create local workspace directory
if (!dir.exists(CLI_WORKSPACE)) {
  dir.create(CLI_WORKSPACE, recursive = TRUE)
  cat("\nCreated local workspace directory:", CLI_WORKSPACE, "\n")
}
```

## Connect and Verify Prerequisites

```{r connect-and-verify}
# Connect to improve repository
improveConnect()

# Verify connection is valid
# This is useful if your session has been idle or if tokens have expired
checkConnect()

setEditable(TRUE)

# Verify tutorial folder from Tutorial 01 exists
tutorialResource <- tryCatch({
  loadResource(TUTORIAL_FOLDER)
}, error = function(e) {
  NULL
})

if (is.null(tutorialResource)) {
  stop(paste("Tutorial folder", TUTORIAL_FOLDER, "does not exist.",
             "Please complete Tutorial 01 first."))
}

cat("Tutorial folder found:", tutorialResource$path, "\n")
cat("Entity ID:", tutorialResource$entityId, "\n")

# Check for expected content from Tutorial 01
children <- loadChildResources(tutorialResource)$data[[1]]
expectedFolders <- c("data", "scripts", "outputs")
foundFolders <- children[children$nodeType == "Folder", "name"]

if (!all(expectedFolders %in% foundFolders)) {
  stop("Expected folders from Tutorial 01 not found. Please complete Tutorial 01 first.")
}

cat("\nFound expected folders from Tutorial 01:\n")
print(foundFolders)
```

## 1. Clone Repository Content

The clone function creates a local copy of repository content for offline work.

```{r clone-content}
cat("\n=== Cloning Repository Content ===\n")



# Clone the entire tutorial folder
cloneResult <- cloneCli(
  localPath = CLI_WORKSPACE,
  ident = tutorialResource$entityId
)

cat("\nClone completed. Local files:\n")
list.files(CLI_WORKSPACE, recursive = TRUE)


```

## 2. Modify Files Locally

Let's modify a file in our local clone.

```{r modify-local}
cat("\n=== Modifying Files Locally ===\n")

# Read the current sample data file
local_data_path <- file.path(CLI_WORKSPACE, "data", "raw", "sample_data.csv")
if (file.exists(local_data_path)) {
  local_data <- read.csv(local_data_path)
  cat("Original data dimensions:", nrow(local_data), "x", ncol(local_data), "\n")
  
  # Modify the data
  local_data$cli_modified <- TRUE
  local_data$timestamp <- Sys.time()
  local_data$new_values <- runif(nrow(local_data))
  
  # Save the modified file
  write.csv(local_data, local_data_path, row.names = FALSE)
  cat("Modified data dimensions:", nrow(local_data), "x", ncol(local_data), "\n")
  cat("Added columns: cli_modified, timestamp, new_values\n")
} else {
  stop("Expected sample_data.csv not found. Please check Tutorial 01 completion.")
}

# Create a new file
new_script_path <- file.path(CLI_WORKSPACE, "scripts", "cli_analysis.R")
new_script_content <- '# Analysis script created via CLI tools
# Created: %s

# Load the modified data
data <- read.csv("../data/raw/sample_data.csv")

# New analysis
cat("Data has CLI modifications:", "cli_modified" %%in%% names(data), "\\n")
cat("Number of records:", nrow(data), "\\n")

# Summary of new values
summary(data$new_values)
'
writeLines(sprintf(new_script_content, Sys.Date()), new_script_path)
cat("\nCreated new script: scripts/cli_analysis.R\n")
```

## 3. Push Changes to Repository

Push the local changes back to the repository.

```{r push-changes}
cat("\n=== Pushing Changes to Repository ===\n")



# Push changes
pushResult <- pushCli(localPath = CLI_WORKSPACE)

cat("\nPush completed. Changes sent to repository.\n")



# Verify changes on server
cat("\n=== Verifying Changes on Server ===\n")
updatedData <- updateResource("./data/raw/sample_data.csv", from = tutorialResource)
cat("Server file updated:", as.character(updatedData$lastModifiedOnDate), "\n")
cat("Version ID:", updatedData$entityVersionId, "\n")

# Check file history
history <- updateHistory(updatedData$entityId)$data[[1]]
cat("\nRecent history:\n")
print(tail(history[, c("entityVersionId", "lastModifiedOnDate", "comment")], 3))
```

## 4. Lock Resources

Locking prevents other users from modifying resources while you're working on them. Since folder locking is currently not working, we'll lock a single file instead.

```{r lock-resources}
cat("\n=== Locking Resources ===\n")

# Load the sample data file we'll be working with
dataFile <- loadResource("./data/raw/sample_data.csv", from = tutorialResource)
cat("File to lock:", dataFile$path, "\n")

# Lock the single file
lockResult <- lockResource(dataFile$entityId)
cat("Locked file:", dataFile$name, "\n")

# Check lock status
lockedFile <- updateResource(dataFile$entityId)
cat("\nLock status:\n")
cat("Locked by:", lockedFile$lockedByName, "\n")
cat("Locked at:", as.character(lockedFile$lockedAt), "\n")

# Store the locked file ID for later unlock
LOCKED_FILE_ID <- lockedFile$entityId
```

## 5. Add Content While Locked

While we have a file locked, let's add a new link via the repository to demonstrate that other operations can still proceed.

```{r add-link-server}
cat("\n=== Adding Link on Server ===\n")

# Get the outputs folder
outputsFolder <- loadResource("./outputs", from = tutorialResource)

# Create a link to the modified data file
newLink <- createLink(
  links =  updatedData$entityId,
  linkName = "cli_modified_data_link",
  linkContainer = outputsFolder$resourceId
)
cat("Created new link:", newLink$path, "\n")
cat("Link points to:", newLink$targetVersionId, "\n")
```

## 6. Pull Changes

Pull the new link to our local workspace.

```{r pull-changes}
cat("\n=== Pulling Changes ===\n")

# Change to workspace directory


# Pull changes
pullResult <- pullCli(localPath = CLI_WORKSPACE)
cat("\nPull completed.\n")

# Check for the new link
local_links <- list.files(file.path(CLI_WORKSPACE,"outputs"), pattern = "*link*")
cat("Links in outputs folder:", paste(local_links, collapse = ", "), "\n")


```

## 7. Delete Link and Push

Remove the link locally and push the deletion.

```{r delete-and-push}
cat("\n=== Deleting Link Locally ===\n")

# Delete the link file
link_path <- file.path(CLI_WORKSPACE, "outputs", "cli_modified_data_link")
if (file.exists(link_path)) {
  unlink(link_path,force = T)
  cat("Deleted local link file\n")
}

# Push the deletion

pushResult <- pushCli(
  localPath = CLI_WORKSPACE
)
cat("\nPush completed. Deletion sent to repository.\n")


# Verify deletion on server
outputsChildren <- updateChildResources(outputsFolder)$data[[1]]
remainingLinks <- outputsChildren[grep("link", outputsChildren$name), "name"]
cat("\nRemaining links in outputs folder:", 
    if(length(remainingLinks) > 0) paste(remainingLinks, collapse = ", ") else "none", "\n")
```

## 8. Unlock Resources

Finally, unlock the file to allow others to modify it.

```{r unlock-resources}
cat("\n=== Unlocking Resources ===\n")

# Unlock the file we locked earlier
if (exists("LOCKED_FILE_ID")) {
  unlockResult <- unlockResource(LOCKED_FILE_ID)
  cat("Unlocked file with ID:", LOCKED_FILE_ID, "\n")
  
  # Verify unlock
  unlockedFile <- updateResource(LOCKED_FILE_ID)
  cat("\nLock status after unlock:\n")
  cat("lockedByName" %in% names(unlockedFile))
  cat("File name:", unlockedFile$name, "\n")
  ""
} else {
  cat("No locked file ID found - skipping unlock\n")
}
```

## Summary

```{r summary}
cat("\n=== Summary of CLI Operations ===\n")

cat("\nOperations completed in this tutorial:\n")
cat("1. ✓ Cloned repository content to local workspace\n")
cat("2. ✓ Modified files locally (data and scripts)\n")
cat("3. ✓ Pushed changes to repository\n")
cat("4. ✓ Locked a single file to prevent concurrent modifications\n")
cat("5. ✓ Added content on server while file was locked\n")
cat("6. ✓ Pulled server changes to local workspace\n")
cat("7. ✓ Deleted content locally and pushed deletion\n")
cat("8. ✓ Unlocked file for others to use\n")

cat("\n=== Key CLI Functions ===\n")
cat("- cliClone(): Create local copy of repository content\n")
cat("- cliPush(): Send local changes to repository\n")
cat("- cliPull(): Get latest changes from repository\n")
cat("- lockResource(): Prevent concurrent modifications\n")
cat("- unlockResource(): Release lock on resources\n")

cat("\n=== Best Practices ===\n")
cat("- Always pull before making changes to avoid conflicts\n")
cat("- Use meaningful commit messages when pushing\n")
cat("- Lock resources when making major changes\n")
cat("- Remember to unlock resources when done\n")
cat("- Keep local workspace organized and clean\n")
```

## Clean Up

```{r cleanup}
# Clean up local workspace (optional)
if (exists("CLI_WORKSPACE") && dir.exists(CLI_WORKSPACE)) {
  # Uncomment to remove workspace:
  # unlink(CLI_WORKSPACE, recursive = TRUE)
  cat("\nLocal workspace preserved at:", CLI_WORKSPACE, "\n")
  cat("You can explore the cloned content there.\n")
}

# Disconnect from improve
improveDisconnect()

cat("\nTutorial completed successfully!\n")
```
