---
title: "CLI 4.5 Features"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CLI 4.5 Features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  eval = FALSE
)
library(improveR)
```

# CLI 4.5 Features

This article covers features introduced in CLI version 4.5.0. For basic CLI
operations (clone, push, pull, lock/unlock), see the [CLI tools](cli-tools.html)
article.

**Requirements**: improveRcontributions with CLI jar 4.5.0 or later.

```{r check-version}
# Verify CLI version
ver <- improveR:::cliDetectedVersion()
cat("CLI version:", ver, "\n")
if (!is.null(ver) && compareVersion(ver, "4.5.0") < 0) {
  stop("This article requires CLI 4.5.0+. Update the JAR in improveRcontributions.")
}
```

## 1. Diff: Compare Local vs Server

The `diffCli()` function shows differences between your local workspace and
the server, similar to `git diff`.

### Unified diff (local vs baseline)

```{r diff-unified}
# After modifying a file locally, see what changed:
diffCli(localPath = "./my-repo", path = "model/control.mod")
```

Output is a standard unified diff:

```
--- a/model/control.mod
+++ b/model/control.mod
@@ -12,3 +12,4 @@
 $ESTIMATION METHOD=1
 $COVARIANCE
+$TABLE ID DV PRED RES
```

### Retrieve specific revisions

Use `--base` or `--theirs` to extract file content at a specific point:

```{r diff-base-theirs}
# Get the baseline (last-pulled) version of a file
baseline <- diffCli("./my-repo", "model/control.mod", base = TRUE)
writeLines(baseline, "control_baseline.mod")

# Get the latest server version (even without pulling)
server <- diffCli("./my-repo", "model/control.mod", theirs = TRUE)
writeLines(server, "control_server.mod")
```

This is useful during conflict resolution: extract both versions, compare them
in your editor, then resolve manually.

## 2. Conflict Resolution Workflow

CLI 4.5 provides a structured workflow for handling concurrent edits.

```{r conflict-workflow}
# Step 1: Check status to see conflicts
statusCli("./my-repo", json = TRUE)

# Step 2: For each conflict file, inspect the versions
# Your local version is already on disk. Get the other two:
diffCli("./my-repo", "model/control.mod", base = TRUE)    # what you both started from
diffCli("./my-repo", "model/control.mod", theirs = TRUE)   # what the other person pushed

# Step 3: See a unified diff of your changes vs baseline
diffCli("./my-repo", "model/control.mod")

# Step 4: Attempt automatic merge
mergeCli("./my-repo", files = "model/control.mod")

# Step 5: If merge succeeded (or after manual edit), mark resolved
resolveCli("./my-repo", files = "model/control.mod")

# Step 6: Push the resolved version
pushCli("./my-repo", comment = "Resolved conflict in control.mod")
```

### Using an external merge tool

```{r merge-tool}
# Use a 3-way merge tool (e.g. kdiff3, meld)
mergeCli("./my-repo",
         files = "model/control.mod",
         mergeTool = "kdiff3 {local} {base} {remote} -o {merged}")
```

## 3. Include/Exclude Patterns for Push

Control which files are uploaded. Patterns use glob syntax (`**`, `*`, `?`,
`[abc]`).

### Setting default patterns

```{r add-inputs}
# Set which files are tracked for push (persisted in the repo)
addInputsCli("./my-repo", includeFiles = "*.R,*.csv,data/")

# Clear and replace patterns
addInputsCli("./my-repo", includeFiles = "model/**,*.mlxtran", clean = TRUE)
```

### Push with patterns

```{r push-patterns}
# Push only specific file types
pushCli("./my-repo",
        comment = "Update R scripts",
        includeFiles = "*.R,*.csv")

# Exclude temporary files
pushCli("./my-repo",
        comment = "Push all except temp",
        excludeFiles = "temp_dir/,*[Tt][Ee][Mm][Pp]*")

# Push specific files by path (like git add + git push)
pushCli("./my-repo",
        comment = "Push two files",
        files = c("model/control.mod", "data/input.csv"))

# Patterns also accept R vectors (collapsed to comma-separated internally)
pushCli("./my-repo",
        comment = "Vector form",
        includeFiles = c("*.R", "data/"))
```

## 4. Reset with Revision

Reset conflict files or the entire workspace to a specific point in time.

```{r reset}
# Reset all conflict files to baseline
resetCli("./my-repo", all = TRUE)

# Reset specific files
resetCli("./my-repo", files = c("model/control.mod", "data/input.csv"))

# Force-reset entire workspace to baseline (destructive)
resetCli("./my-repo", force = TRUE, confirm = TRUE)

# Reset to a specific revision ID
resetCli("./my-repo", force = TRUE, confirm = TRUE,
         revision = "D504E3C0BEE345278A90359E1DEA1CBB")
```

## 5. Reset Inventory and Files-to-Keep

When re-running a step, you can control what happens to previous outputs.

```{r reset-inventory}
# Normal rerun: previous outputs remain, new outputs are added
runStepResource("/Projects/analysis/Step 1")

# Reset inventory: wipe all outputs before running
# The step starts with a clean inventory
runStepResource("/Projects/analysis/Step 1", resetInventory = TRUE)

# Reset inventory but keep specific files (by resource GUID)
# Useful when you want to preserve certain outputs across reruns
inventory <- loadFullChildResources("/Projects/analysis/Step 1")$data[[1]]
keepIds <- inventory$resourceId[inventory$name %in% c("important_result.rds")]
runStepResource("/Projects/analysis/Step 1",
                resetInventory = TRUE,
                filesToKeep = keepIds)
```

## 6. Update Process Properties on Live Steps

Set arbitrary process properties on an already-realised step without losing
existing configuration.

```{r update-process-property}
# Set checkout include patterns on a live step
updateProcessProperty(
  stepIdent = "/Projects/analysis/Step 1",
  processName = "Main",
  key = "checkoutIncludePatterns",
  value = "test_script.R\r\ndata.csv"
)

# This is a safe read-modify-write operation:
# existing properties (toolId, runserverToolId, etc.) are preserved
```

For setting properties at step creation time, use `setProcessValue` on the
template before `realise()`:

```{r set-process-value-template}
template <- createStepTemplateEnv(treeIdent = myTree)
template$setStepRunserverLabel("runserver")
template$setStepToolLabel("Rbatch")
template$setStepToolInstance("default")

# Set checkout patterns before realising
template$setProcessValue("Main", "checkoutIncludePatterns", "*.R\r\n*.csv")

template$realise()
```

## 7. Relation Type Management

Create and delete relation types (admin operation).

```{r relation-types}
# List existing relation types
types <- loadRelationTypes()
print(types[, c("id", "name", "reverseName")])

# Create a new relation type
newType <- createRelationType(
  name = "derived from",
  reverseName = "source of",
  description = "Links a derived dataset to its source"
)
cat("Created type:", newType$id, "\n")

# Use it to create a relation between resources
createResourceRelation(
  ident = "/Projects/analysis/derived_data.csv",
  targetIdent = "/Projects/data/raw_data.csv",
  relationTypeId = newType$id,
  description = "Derived via cleaning script"
)

# Delete a relation type (when no longer needed)
deleteRelationType(newType$id)
```

## 8. Repo Prefix Auto-Discovery

When connecting without `IMPROVER_STEP`, the repository entity ID prefix
is now auto-discovered:

```{r repo-prefix}
# Connect without IMPROVER_STEP set
Sys.setenv(IMPROVER_STEP = "")
improveConnect()

# repoPrefix() now returns the correct prefix (e.g. "hc4310:")
# instead of "" — all entity ID operations work correctly
cat("Prefix:", repoPrefix(), "\n")
```

## Summary of New Functions

| Function | Description |
|---|---|
| `diffCli()` | Unified diff or revision content streaming |
| `resetCli(revision = ...)` | Reset to specific revision |
| `addInputsCli()` | Set default include patterns |
| `pushCli(includeFiles = ...)` | Push with file patterns (vector or string) |
| `runStepResource(resetInventory, filesToKeep)` | Control inventory on rerun |
| `updateProcessProperty()` | Set process properties on live steps |
| `createRelationType()` | Create relation types (admin) |
| `deleteRelationType()` | Delete relation types (admin) |
| `repoPrefix()` | Get repository entity ID prefix |
