---
title: "Getting started with {improveR}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with improveR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

**improveR** is the R interface to Scinteco's improve platform - a secure, versioned repository for pharmaceutical R&D workflows. It provides seamless integration between R and the improve platform, enabling 21 CFR Part 11 compliant workflow management for pharmaceutical modeling, biostatistical analysis, and data science.

Key capabilities include:

- **Workflow Management**: Create and execute multi-step pharmaceutical modeling workflows
- **Version Control**: Full versioning of data, code, and analyses with complete audit trails
- **Resource Management**: Load and save analytical resources with metadata and lineage tracking
- **Collaborative Research**: Team-based review system for pharmaceutical research

## Installation

Install improveR from the Scinteco repository:

```{r install}
# Add the improve repository
options(repos = c(getOption("repos"),
  improve = "https://r.scinteco.com/4.4.0/"))

# Install the package
install.packages("improveR")

# Optional: Install CLI tools
install.packages("improveRcontributions")
```

For detailed installation instructions, see `vignette("install-improveR")`.

## Setup and Connection

Configure your environment variables to connect to the improve repository:

```{r setup-env}
# Add to your .Renviron file:
# Sys.setenv(IMPROVER_REPO_URL = "https://<url>:<port>/repository")
# Sys.setenv(IMPROVER_STEP = "<valid-entityId>")

# Load the package
library(improveR)

# Connect to the repository
improveConnect()

# Enable editing
setEditable(TRUE)

# Verify connection
whoami()
```

## Core Concepts
### Resources
The improve platform organizes everything as **resources** - versioned entities with unique identifiers:

- **Folders**: Organize resources hierarchically
- **Files**: Store data, scripts, and outputs with automatic versioning
- **Links**: Create stable references to specific versions of resources
- **Steps**: Executable units of analytical work within workflows

```{r basic-operations}
# Load a resource by path
resource <- loadResource("/Projects/my-analysis/data")

# Load by entity ID
resource <- loadResource("EN-12345")

# List folder contents
children <- loadChildResources(resource)

# Create a new file
myData <- data.frame(x = 1:10, y = rnorm(10))
write.csv(myData, "data.csv", row.names = FALSE)

dataFile <- createFile(
  targetIdent = resource$entityId,
  fileName = "data.csv",
  localPath = "data.csv",
  comment = "My analysis data"
)
```

### Workflows

Workflows chain together analytical **steps** organized in **analysis trees**:

```{r workflow-basics}
# Create an analysis tree
analysisTree <- createAnalysisTree(
  targetIdent = "/Projects/my-analysis",
  treeName = "EDA_Workflow",
  comment = "Exploratory data analysis"
)

# Create a step environment
stepEnv <- createStepTemplateEnv(treeIdent = analysisTree)

# Configure the step
stepEnv$setStepDescription("Load and prepare data")
stepEnv$setStepRationale("Initial data loading and validation")

# Add input files
stepEnv$addStepRemoteFile(ident = dataFile$entityId, variableName = "input-data")

# Set execution configuration
stepEnv$setStepRunserverLabel("runserver")
stepEnv$setStepToolLabel("R_4.2")
stepEnv$setStepToolInstance("rbatch")

# Create the step
resultStep <- stepEnv$realise()
step <- resultStep$getStepResource()
```

### Navigating Workflows

Use `getStep()` to explore workflow relationships:

```{r workflow-navigation}
# Get a step with its full context
stepEnv <- getStep("/Projects/my-analysis/EDA_Workflow/Step 1")

# Load and explore child steps
stepEnv$children$load()
ls(stepEnv$children)

# Load and explore lineage (upstream dependencies)
stepEnv$lineage$load(treeDepth = -1)
workflow <- stepEnv$workflow$df()

# Load usage (downstream consumers)
stepEnv$usage$load(treeDepth = -1)

# Check step state
stepEnv$getStepState()

# Get step inventory
inventory <- stepEnv$getStepInventory()
```

## Next Steps

Now that you understand the basics, explore these topics in more detail:

- **[Basic Node Types](basic-node-types.html)**: Detailed guide to folders, files, links, and versioning
- **[Workflows Basic](workflows-basic.html)**: Create and execute your first workflow
- **[Working with Steps](workingWithSteps.html)**: Advanced step management and navigation
- **[Working with Workflows](workingWithWorkflows.html)**: Workflow orchestration and execution
- **[CLI Tools](cli-tools.html)**: Git-like version control operations

## Getting Help

```{r help}
# Function documentation
?improveConnect
?createFile
?getStep

# List all improveR functions
ls("package:improveR")

# Disconnect when done
improveDisconnect()
```

For support, visit the [Scinteco website](https://www.scinteco.com/).
