---
title: "Install improveR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Install improveR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo=TRUE,
  message=TRUE,
  warning=TRUE,
  eval=FALSE
)
library(improveR)
```

# Installing improveR

This guide will walk you through installing the improveR package from the improve repository.

## Step 1: Configure R Repository

First, we need to add the improve repository as a secondary repository source:

```{r configure-repo}
# Add improve repository
options(repos = c(getOption("repos"),
  improve = "https://r.scinteco.com/4.4.0/"))

# Verify the configuration
cat("\nUpdated repositories:\n")
print(getOption("repos"))
```

## Step 2: Install improveR Package

Now we can install the improveR package from the improve repository:

```{r install-improver}
# Install improveR package
install.packages("improveR")
#this package contains the CLI tools
install.packages("improveRcontributions")
# Load the package to verify installation
library(improveR)

# Check package version
cat("improveR package version:", as.character(packageVersion("improveR")), "\n")
```

## Step 3: Configure Environment Variables

To use improveR, you need to set up the required environment variables. 
We recommend creating a `.Renviron` file in your project root:

```{r env-vars-display, eval=FALSE}
# Add these lines to your .Renviron file:
# Sys.setenv(IMPROVER_REPO_URL = "https://<url>:<repoPort>/repository")
# Sys.setenv(IMPROVER_STEP = "<valid-entityId>")  # Ideally the entityId of your working folder
```

## Step 4: Verify Installation

Let's verify that improveR is properly installed and configured:

```{r verify-installation}
# Check if improveR is installed
if ("improveR" %in% installed.packages()[, "Package"]) {
  cat("✓ improveR package is installed\n")
  
  # List key functions
  cat("\nKey improveR functions available:\n")
  improver_functions <- ls("package:improveR")
  main_functions <- improver_functions[improver_functions %in% c(
    "improveConnect", "improveDisconnect", "loadResource", "createFile", 
    "createFolder", "createLink", "effectiveRights", "whoami"
  )]
  print(main_functions)
} else {
  cat("✗ improveR package is NOT installed\n")
}

# Check environment variables
cat("\nEnvironment variable status:\n")
repo_url <- Sys.getenv("IMPROVER_REPO_URL")
step_id <- Sys.getenv("IMPROVER_STEP")

if (repo_url != "") {
  cat("✓ IMPROVER_REPO_URL is set:", repo_url, "\n")
} else {
  cat("✗ IMPROVER_REPO_URL is NOT set\n")
}

if (step_id != "") {
  cat("✓ IMPROVER_STEP is set:", step_id, "\n")
} else {
  cat("✗ IMPROVER_STEP is NOT set\n")
}
```

## Step 5: Test Connection (Optional)

If you have valid environment variables set, you can test the connection:

```{r test-connection, eval=FALSE}
# Only run this if you have valid credentials
if (Sys.getenv("IMPROVER_REPO_URL") != "" && Sys.getenv("IMPROVER_STEP") != "") {
  tryCatch({
    # Connect to repository
    improveConnect()
    
    # Test basic functionality
    user <- whoami()
    cat("\nSuccessfully connected as:", user, "\n")
    
    # Disconnect
    improveDisconnect()
    cat("Connection test completed successfully!\n")
  }, error = function(e) {
    cat("Connection test failed:\n")
    print(e)
  })
} else {
  cat("\nSkipping connection test - environment variables not set\n")
  cat("Please configure IMPROVER_REPO_URL and IMPROVER_STEP before testing connection\n")
}
```

## Troubleshooting

If you encounter issues:

1. **Package installation fails:**
   - Check your internet connection
   - Verify the repository URL is accessible
   - Try installing with `type = "source"` if binary installation fails

2. **Environment variables not working:**
   - Restart your R session after creating `.Renviron`
   - Check that `.Renviron` is in your working directory
   - Use `Sys.setenv()` for temporary testing

3. **Connection issues:**
   - Verify your repository URL format (https://<url>:<port>/repository)
   - Check that your step ID is valid
   - Ensure you have network access to the repository

## Summary

You have now:
1. ✓ Added the improve repository to your R configuration
2. ✓ Installed the improveR package
3. ✓ Learned how to configure environment variables
4. ✓ Verified the installation

Next steps:
- Configure your environment variables in `.Renviron`
- Try the basic node types tutorial: `tutorial-01-basic-node-types.Rmd`
- Explore the improveR documentation

```{r session-info}
# Display session information for debugging
cat("\n=== Session Information ===\n")
sessionInfo()
```
