---
title: "Workflows basic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflows basic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo=TRUE,
  message=TRUE,
  warning=TRUE,
  eval=FALSE
)
library(improveR)
```

# Tutorial 03: Basic Workflows

This tutorial introduces the workflow concepts in improve by creating a simple step that loads and processes an R dataset. You'll learn how to create analysis trees, configure steps, and execute them.

**Prerequisites**: Complete Tutorials 01 and 02 first, as this tutorial builds on those concepts.

## Environment Variables

This tutorial requires the following environment variables:

```{r env-vars-display, eval=FALSE}
# Add these lines to your .Renviron file:
# Sys.setenv(IMPROVER_REPO_URL = "https://<url>:<repoPort>/repository")
# Sys.setenv(IMPROVER_STEP = "<valid-entityId>")  # Ideally the entityId of your tutorial folder
```

## Tutorial Configuration

```{r tutorial-config}
# Define paths and configuration
TUTORIAL_FOLDER <- "/Projects/tutorials/myUserName/tutorial3-workflows"  # Update with your username

# Configure runserver settings as variables (not environment variables)
# These would typically come from your organization's configuration
R_RUNSERVER <- "runserver"  # Or your specific runserver name
R_TOOL <- "R_4.2"       # R version specification
R_TOOL_INSTANCE <- "rbatch"  # Instance type for R execution

# Check environment variables
if (Sys.getenv("IMPROVER_REPO_URL") == "") {
  stop("IMPROVER_REPO_URL environment variable is not set.")
}

if (Sys.getenv("IMPROVER_STEP") == "") {
  stop("IMPROVER_STEP environment variable is not set.")
}

cat("Repository URL:", Sys.getenv("IMPROVER_REPO_URL"), "\n")
cat("Tutorial folder:", TUTORIAL_FOLDER, "\n")
cat("R Runserver:", R_RUNSERVER, "\n")
cat("R Tool:", R_TOOL, "\n")
cat("R Tool Instance:", R_TOOL_INSTANCE, "\n")
```

## Connect and Setup

```{r connect-setup}
# Connect to improve repository
improveConnect()

# Verify connection is valid
# This is useful if your session has been idle or if tokens have expired
checkConnect()

# Display repository version
repoVersion <- getRepositoryVersion()
if (!is.null(repoVersion)) {
  cat("Connected to repository version:", repoVersion, "\n")
}

setEditable(TRUE)

# Verify tutorial folder exists
tutorialResource <- tryCatch({
  loadResource(TUTORIAL_FOLDER)
}, error = function(e) {
  NULL
})

if (is.null(tutorialResource)) {
  # Create tutorial folder if it doesn't exist
  parentPath <- dirname(TUTORIAL_FOLDER)
  folderName <- basename(TUTORIAL_FOLDER)
  
  parent <- loadResource(parentPath)
  if (is.null(parent)) {
    stop(paste("Parent folder", parentPath, "does not exist. Please create it first."))
  }
  
  tutorialResource <- createFolder(
    folderName = folderName,
    targetIdent = parent$entityId,
    comment = "Tutorial 03: Workflows"
  )
  cat("Created tutorial folder:", tutorialResource$path, "\n")
} else {
  cat("Tutorial folder found:", tutorialResource$path, "\n")
}
```

## 1. Create Script Files

First, let's create R scripts that our workflow steps will execute.

```{r create-scripts}
cat("\n=== Creating Script Files ===\n")

# Create a simple data loading script
dataLoadScript <- '# Data Loading Script
# This script loads the built-in iris dataset and saves it as CSV

# Load the iris dataset
data(iris)

# Display basic information
cat("Dataset: iris\\n")
cat("Dimensions:", nrow(iris), "rows x", ncol(iris), "columns\\n")
cat("Column names:", paste(names(iris), collapse = ", "), "\\n")

# Summary statistics
summary(iris)

# Save to CSV
write.csv(iris, "iris_data.csv", row.names = FALSE)
cat("\\nData saved to iris_data.csv\\n")
'

# Write script locally
writeLines(dataLoadScript, "load_iris_data.R")

# Upload to repository
scriptFile <- createFile(
  targetIdent = tutorialResource$entityId,
  fileName = "load_iris_data.R",
  localPath = "load_iris_data.R",
  comment = "Script to load and save iris dataset"
)
cat("Uploaded script:", scriptFile$path, "\n")

# Create an analysis script
analysisScript <- '# Data Analysis Script
# This script performs basic analysis on the iris dataset

# Check if data file exists
if (!file.exists("iris_data.csv")) {
  stop("iris_data.csv not found. Run data loading step first.")
}

# Load the data
iris_data <- read.csv("iris_data.csv")
cat("Loaded data with", nrow(iris_data), "observations\\n")

# Create visualizations
pdf("iris_analysis.pdf")

# Pairwise scatter plots
pairs(iris_data[, 1:4], 
      main = "Iris Dataset - Pairwise Relationships",
      pch = 21, 
      bg = c("red", "green", "blue")[unclass(iris_data$Species)])

# Box plots by species
par(mfrow = c(2, 2))
for (col in names(iris_data)[1:4]) {
  boxplot(iris_data[[col]] ~ iris_data$Species,
          main = col,
          xlab = "Species",
          ylab = col,
          col = c("red", "green", "blue"))
}

dev.off()
cat("\\nAnalysis plots saved to iris_analysis.pdf\\n")

# Calculate means by species
species_means <- aggregate(. ~ Species, data = iris_data, FUN = mean)
write.csv(species_means, "species_means.csv", row.names = FALSE)
cat("Species means saved to species_means.csv\\n")

# Print summary
print(species_means)
'

writeLines(analysisScript, "analyze_iris_data.R")

analysisFile <- createFile(
  targetIdent = tutorialResource$entityId,
  fileName = "analyze_iris_data.R",
  localPath = "analyze_iris_data.R",
  comment = "Script to analyze iris dataset"
)
cat("Uploaded analysis script:", analysisFile$path, "\n")

# Clean up local files
unlink(c("load_iris_data.R", "analyze_iris_data.R"))
```

## 2. Create Analysis Tree

Analysis trees organize workflow steps hierarchically.

```{r create-tree}
cat("\n=== Creating Analysis Tree ===\n")

# Create an analysis tree
analysisTree <- createAnalysisTree(
  targetIdent = tutorialResource$entityId,
  treeName = "Iris_Analysis_Workflow",
  comment = "Workflow for loading and analyzing iris dataset"
)

cat("Created analysis tree:", analysisTree$name, "\n")
cat("Tree ID:", analysisTree$entityId, "\n")
```

## 3. Create Data Loading Step

Now we'll create our first workflow step using the step template environment.

```{r create-load-step}
cat("\n=== Creating Data Loading Step ===\n")

# Create step environment
loadStepEnv <- createStepTemplateEnv(treeIdent = analysisTree)

# Configure the step
loadStepEnv$setStepDescription("Load Iris Dataset")
loadStepEnv$setStepRationale("Load the built-in iris dataset and save as CSV for further analysis")

# Set runserver configuration
loadStepEnv$setStepRunserverLabel(R_RUNSERVER)
loadStepEnv$setStepToolLabel(R_TOOL)
loadStepEnv$setStepToolInstance(R_TOOL_INSTANCE)

# Add the script file as the command file
loadStepEnv$addStepRemoteFile(ident = scriptFile$entityId, 
  variableName = "command-file"
)

# Realize the step (create it in the repository)
loadResultStep <- loadStepEnv$realise()
loadStepEnv$finishRun()
# Get the created step resource
loadStep <- loadResultStep$getStepResource()
cat("Created step:", loadStep$name, "\n")
cat("Step path:", loadStep$path, "\n")
cat("Step ID:", loadStep$entityId, "\n")
dataFile <- dplyr::filter(loadResultStep$getStepInventory()$data[[1]],name=="iris_data.csv")
```

## 4. Create Analysis Step

Create a second step that depends on the output of the first step.

```{r create-analysis-step}
cat("\n=== Creating Analysis Step ===\n")

# Create step environment for analysis
analysisStepEnv <- createStepTemplateEnv(treeIdent = analysisTree)

# Configure the step
analysisStepEnv$setStepDescription("Analyze Iris Dataset")
analysisStepEnv$setStepRationale("Perform exploratory data analysis on the iris dataset")

# would Set this as a child of the load step
# analysisStepEnv$setStepParent(loadStep$resourceId)

# Set runserver configuration
analysisStepEnv$setStepRunserverLabel(R_RUNSERVER)
analysisStepEnv$setStepToolLabel(R_TOOL)
analysisStepEnv$setStepToolInstance(R_TOOL_INSTANCE)

# Add the analysis script
analysisStepEnv$addStepRemoteFile(
  analysisFile$entityId,
  variableName = "command-file"
)

# We'll also link the expected output from the parent step
analysisStepEnv$addStepRemoteFile(
  dataFile
)

# Realize the step
analysisResult <- analysisStepEnv$realise()

# Get the created step
analysisStep <- analysisResult$getStepResource()
cat("Created analysis step:", analysisStep$name, "\n")
cat("Step path:", analysisStep$path, "\n")
cat("Parent step:", loadStep$name, "\n")
```

## 5. Explore Step Structure

Let's examine the structure of our workflow.

```{r explore-structure}


# Get step inventory (files in the step)
cat("\n=== Step Inventory ===\n")

# For load step
loadInventory <- loadChildResources(loadStep)$data[[1]]
cat("\nLoad step contains:\n")
print(loadInventory[, c("name", "nodeType")])

# For analysis step
analysisInventory <- loadChildResources(analysisStep)$data[[1]]
cat("\nAnalysis step contains:\n")
print(analysisInventory[, c("name", "nodeType")])
```



## 8. Workflow Best Practices

```{r best-practices}
cat("\n=== Workflow Best Practices ===\n")

cat("
1. Step Organization:
   - Use descriptive names for steps
   - Document rationale for each step
   - Organize steps hierarchically in analysis trees

2. Data Flow:
   - Use links to connect step outputs to inputs
   - Links ensure reproducibility by pointing to specific versions
   - Document expected inputs and outputs

3. Configuration:
   - Store runserver configuration in variables or config files
   - Use consistent tool versions across related steps
   - Document tool requirements

4. Error Handling:
   - Check for required input files
   - Provide meaningful error messages
   - Save logs and intermediate results

5. Reproducibility:
   - All inputs should be versioned
   - Record tool versions and parameters
   - Capture execution environment details
")
```

## Summary

```{r summary}
cat("\n=== Summary ===\n")

# List all created resources
allResources <- loadChildResources(tutorialResource)$data[[1]]
workflowResources <- allResources[allResources$nodeType %in% c( "Step"), ]

cat("Created workflow resources:\n")
print(workflowResources[, c("name", "nodeType", "path")])

cat("\n=== Key Concepts Covered ===\n")
cat("1. Analysis Trees - Organize workflow steps\n")
cat("2. Steps - Executable units of work\n")
cat("3. Step Configuration - Runserver, tools, and instances\n")
cat("4. Step Dependencies - Parent-child relationships\n")
cat("5. Data Flow - Using links between steps\n")

cat("\n=== Next Steps ===\n")
cat("- In production, steps are executed by the configured runserver\n")
cat("- Use runStep() to submit steps for execution\n")
cat("- Monitor step execution status and logs\n")
cat("- Build more complex workflows with multiple branches\n")
```

## Clean Up

```{r cleanup}


cat("\nTutorial completed successfully!\n")
cat("Workflow created in:", TUTORIAL_FOLDER, "\n")
cat("\nNote: This tutorial demonstrated workflow concepts.\n")
cat("In production use, steps would be executed by the runserver.\n")
```
