---
title: "How to..."
format:
  html:
    toc: true
    toc-depth: 4
    toc-title: Contents
    toc-location: left
    #toc-collapsed: false
    code-fold: show
    number-sections: true
    number-depth: 2
    # smooth-scroll: true
    # message: false
    # warning: false
    # highlight-style: nord
    # output-file: 
    embed-resources: true
    # theme: 
    #   light: [materia, style.scss]
    #   dark: slate
    fontsize: 16px
    # grid:
    # sidebar-width: 300px
    fig-cap-location: top
    fig-align: left
    code-annotations: true
knitr:
    opts_chunk: 
      collapse: true
      dev: "ragg_png"
# params:
#   include_questions: true
editor_options: 
  markdown: 
    wrap: 72
# project:
#   execute-dir: project
lightbox: true
---


```{r}
#| label: setup
library(improveR)
```


# ... all steps of an analysis Tree

```{r}
#| eval: false

childResourcesATModeling <- improveR::loadChildResources(ident='/improve-tutorial/Modeling/') %>%
  select(data) %>%
  unnest(data) %>%
  select(""name"", ""entityId"", ""path"")

```


# ... get the workflow leading to a step

```{r}
#| eval: false
envATReportingStep1 <- getStep(""/improve-tutorial/Reporting/Step 1"")

# get the workflow and show details on Step 'Modeling/Step 10'
envATReportingStep1$lineage$load(treeDepth=-1)

#show all steps which form part of the workflow
envATReportingStep1$workflow$df()
```

Note that before calling `lineage$load()`, the `df()` in the `workflow` binding will only return the step itself, i.e.
`Reporting/Step 1`.


# ... get the elements aka resources of a step's inventory

```{r}
#| eval: false

# appraoch 1 
stepInventory <- loadChildResources(ident="/improve-tutorial/Modeling/Step 1")$data[[1]]
ncol(stepInventory)

# approach 2
envStep <- getStep(ident="/improve-tutorial/Modeling/Step 1")
stepInventory2 <- envStep$getStepInventory()$data[[1]]
ncol(stepInventory2)

# the approach with getStep returns two additional columns
setdiff(names(stepInventory2), names(stepInventory)) # 'isVersion', 'inventoryPath'

# approach 3
envStep <- getStep(ident="/improve-tutorial/Modeling/Step 1")
envStep$stepDf #no does not show content of inventory; in the inventory is more than remote files;



```
# ... remove a step from a workflow

```{r}
#| eval: false
#| code-summary: "Remove step from workflow leading a specific step"
step$workflow$removeStep()
```

# ... get the outdated elements of an inventory

```{r}
envATReportingStep1 <- getStep("/improve-tutorial/Reporting/Step 1")
changedAndOutdatedFiles <- envATReportingStep1$workflow$changedAndOutdatedFiles()

#outdated elements in an inventory
inventory <- envATReportingStep1$getStepInventory()
inventoryOutdatedLinks <- inventory %>%
  select(data) %>%
  unnest(data) %>%
  filter(outdatedLink==TRUE) 
nrow(inventoryOutdatedLinks)

#outdated link in resource
x <- loadResource("/improve-tutorial/Reporting/Step 1")
x$outdatedLink  #outdatedLink can be FALSE even if in the inventory of the Step are outdated elements

```

* elaborate on the difference between outdated elements of an inventory and changedAndOutdated in the workflow context
* outdatedAndChanged is different; also includes changed; changes input is a different one; outdatedLink=>here only outdated links; changed: files are not the same; changed red frame; outdated yellow frame



# ... rerun changed and outdated steps with in a workflow

```{r}
#| eval: false



step$workflow$rerunChangedandOutdated
```


# Common Errors

## Tooling

When defining a step's tooling, no interactive e.g. RStudio instance is permissble. Executing a step via {improveR} requires a tooling
which can be run in batch mode. 

