---
title: "How to..."
format:
  html:
    toc: true
    toc-depth: 1
    toc-title: Contents
    toc-location: right
    code-fold: false
    number-sections: false
    number-depth: 2
    # smooth-scroll: true
    # message: false
    # warning: false
    # highlight-style: nord
    # output-file:
    # embed-resources: true
    # theme: 
    #   light: [materia, style.scss]
    #   dark: slate
    fontsize: 16px
    # grid:
    # sidebar-width: 300px
    fig-cap-location: top
    fig-align: left
    code-annotations: true
knitr:
    opts_chunk: 
      collapse: true
      dev: "ragg_png"
# params:
#   include_questions: true
editor_options: 
  markdown: 
    wrap: 72
# project:
#   execute-dir: project
lightbox: true
---

```{r}
#| include: false 
library(improveR)
library(tidyverse, quietly = T)
library(tictoc) #time keeping of functions
options(width=10000) #make results window wide; horizontal scrollbar


fn_connect <- function() {
  Sys.setenv(
    IMPROVER_REPO_URL = "http://envhost1.hc.scintecodev.internal:5310/repository"
    
  )
  Sys.setenv(IMPROVER_AUTO_AUTH = TRUE)

  clearConnectionData()
  #improveRtestsupport::improveConnect()
  improveR::improveConnect()

  setEditable(TRUE)
  print(Sys.getenv("IMPROVER_REPO_URL"))
}

fn_connect()

improveConnected()
```

# Common Tasks

## and how to solve them

This guide provides quick examples for common {improveR} operations.

# attach a step to a parent

See [reference](../reference/attachStep.html). See [article](../articles/workingWithSteps.html#attachstep).

```{r}
#| eval: false
attachStep(ident, parent, from = pwd())
```

# change the description of a step

See [reference](../reference/changeStepDescription.html). See [article](../articles/workingWithSteps.html#changestepdescription).

```{r}
#| eval: false
changeStepDescription(ident=stepIdent, description="new description")
```

# change the rationale of a step

See [reference](../reference/changeStepRationale.html). See [article](../articles/workingWithSteps.html#changesteprationale).

```{r}
#| eval: false
changeStepRationale(ident=stepIdent, description="new description")
```

# create a step environment

See [reference](../reference/createStepEnv.html). See [article](../articles/workingWithSteps.html#createstepenv).
```{r}
#| eval: false
stepDf <- improveR:::getStepDf(stepIdent)
workflow <- createWorkflow()
createStepEnv(stepDf, workflow)
```

# create an analyis tree

See [reference](../reference/createAnalysisTree.html). See [article](../articles/workingWithSteps.html#realise-step).
```{r}
#| eval: false
createAnalysisTree(targetIdent="/improve-tutorial/", treeName="newTree", comment = "this is a comment")
```

# create a folder

See [reference](../reference/createFolder.html). See [article](../articles/workingWithSteps.html#specify-destination).
```{r}
#| eval: false
createFolder(folderName="NameOfNewFolder", targetIdent="/improve-tutorial/") 
```


# create an external Link

See [reference](../reference/createExternalLink.html). See [article](../articles/basic-node-types.html#working-with-external-links).

```{r}
#| eval: false
createExternalLink(targetIdent="/improve-tutorial/demoSteps/Step 1", linkName="Weblink to scinteco", url="https://www.scinteco.com/", comment="Resource to integrate.")

```


# create a step template environment

See [reference](../reference/createStepTemplateEnv.html). See [article](../articles/workingWithSteps.html#createsteptemplateenv).
```{r}
#| eval: false
stepDf <- improveR:::getStepDf(stepIdent)
workflow <- createWorkflow()
createStepEnv(stepDf, workflow)
```


# export a workflow

See [reference](../reference/exportWorkflow.html). See [article](../articles/workingWithWorkflows.html#exporting-and-importing-workflows).

```{r}
#| eval: false

stepEnv <- getStep("/improve-tutorial/Reporting/Step 1") #get environment of step to which the workflow leads
stepEnv$dependencies$load(treeDepth=-1, stepDepth=-1) #get the full dependencies with -1
exportWorkflow(envATReportingStep1$workflow, "WFReportingSTep1") #export the workflow binding of stepEnv via exportWorfklow
```


# get all steps of an analysis tree

```{r}
#| eval: false
#| code-fold: false

childResourcesATModeling <- improveR::loadChildResources(ident='/improve-tutorial/Modeling/') %>%
  select(data) %>%
  unnest(data) %>%
  select(""name"", ""entityId"", ""path"")
```

# get all analysis trees within a folder

Get all analysis trees inside the folder 'improve-tutorial'.

```{r}
#| cache: true
allAnalysisTrees <- improveR::loadChildResources(ident="/improve-tutorial/")
allAnalysisTrees$data[[1]] %>% select(nodeType, name, path)
```

# get all child steps of a step

## Version 1: loadChildSteps()
```{r}
#| code-fold: false
dfChild <- loadChildSteps(ident="/improve-tutorial/Modeling/Step 1")
dfChild$data[[1]] %>%
  select(path, resourceId)
```

## Version 2: getStep()
```{r}
#| code-fold: false
envStep <- getStep(ident="/improve-tutorial/Modeling/Step 1")
envStep$children$load()
ls(envStep$children)
```

# get all available tools

See [reference](../reference/getToolInstances.html).

```{r}
#| code-fold: false
#| eval: false

envTools<-getToolInstances()
ls(envTools)

# [1] "Nonmem 7.4 7.4 runserver"                                                                 
# [2] "Nonmem 7.4 7.4 slurm runserver"                                                           
# [3] "Nonmem nm4robot_desktop-pmalkg2_cflandorfer nmLsf_desktop-pmalkg2_cflandorfer runserver"  
# [4] "Nonmem nm4robot_desktop-pmalkg2_cflandorfer nmSlurm_desktop-pmalkg2_cflandorfer runserver"
# [5] "R R4.2 R4.2 runserver"        

```

# get the workflow leading to a step
```{r}
#| code-fold: false
#| eval: false
envATReportingStep1 <- getStep("/improve-tutorial/Reporting/Step 1")

# get the workflow and show details on Step 'Modeling/Step 10'
envATReportingStep1$dependencies$load(treeDepth=-1)

#show all steps which form part of the workflow
envATReportingStep1$workflow$df()
```

Note that before calling `dependencies$load()`, the `df()` in the `workflow` binding will only return the step itself, i.e.,
`Reporting/Step 1`.


# get the elements aka resources of a step's inventory

```{r}
#| code-fold: false
#| eval: false

# approach 1 
stepInventory <- loadChildResources(ident="/improve-tutorial/Modeling/Step 1")$data[[1]]
ncol(stepInventory)

# approach 2
envStep <- getStep(ident="/improve-tutorial/Modeling/Step 1")
stepInventory2 <- envStep$getStepInventory()$data[[1]]
ncol(stepInventory2)

# the approach with getStep returns two additional columns
setdiff(names(stepInventory2), names(stepInventory)) # 'isVersion', 'inventoryPath'

# approach 3
envStep <- getStep(ident="/improve-tutorial/Modeling/Step 1")
envStep$stepDf #does not show content of inventory; the inventory contains more than remote files
```

# remove a step from a workflow

```{r}
#| eval: false
#| code-fold: false
#| code-summary: "Remove step from workflow leading a specific step"
step$workflow$removeStep()
```

# get the outdated elements of an inventory

```{r}
#| eval: false
#| code-fold: false

envATReportingStep1 <- getStep("/improve-tutorial/Reporting/Step 1")
changedAndOutdatedFiles <- envATReportingStep1$workflow$changedAndOutdatedFiles()

#outdated elements in an inventory
inventory <- envATReportingStep1$getStepInventory()
inventoryOutdatedLinks <- inventory %>%
  select(data) %>%
  unnest(data) %>%
  filter(outdatedLink==TRUE) 
nrow(inventoryOutdatedLinks)

#outdated link in resource
x <- loadResource("/improve-tutorial/Reporting/Step 1")
x$outdatedLink  #outdatedLink can be FALSE even if there are outdated elements in the inventory of the Step

```

<!-- * Elaborate on the difference between outdated elements of an inventory and 'changedAndOutdated' in the workflow context.
* 'outdatedAndChanged' is different from just 'outdated'. It also includes files that have changed.
* 'outdatedLink' refers only to outdated links (yellow frame).
* 'changed' refers to files that are not the same (red frame). -->

# rerun changed and outdated steps within a workflow

```{r}
#| eval: false
#| code-fold: false
step$workflow$rerunChangedAndOutdated()
```

# create a step template

```{r}
#| eval: false
#| code-fold: false
stepTemplateEnv <- createStepTemplateEnv()
```

# delete a resource

See [reference](../reference/delete.html).

```{r}
#| include: false
#| code-summary: "setup analysis tree with only one step; which is deleted below"
createAnalysisTree(targetIdent = "/improve-tutorial/", treeName="yourTree")
improveR:::createStep(treeIdent = "/improve-tutorial/yourTree")
```

```{r}
#| code-summary: "Delete step from analysis tree"

#Get steps inside the analysis tree
steps <- loadChildResources("/improve-tutorial/yourTree")$data[[1]] 
steps %>% select(nodeType, name, path, entityId)

#Delete the steps via their entityId, or equivalent ident
delete(res=steps$entityId)

#Loading child resources of tree again confirms that step(s) were removed
loadChildResources("/improve-tutorial/yourTree")$data[[1]]
```

```{r}
#| include: false
#| code-summary: "Reset folder"
delete(res="/improve-tutorial/yourTree")
```

# terminate the execution of a step

See [reference](../reference/terminateStepResource.html). See [article](../articles/workingWithSteps.html).

```{r}
#| eval: false
terminateStepResource(ident="/improve-tutorial/Modeling/Step 9")
```

# trigger the execution of a step
```{r}
#| eval: false
runStepResource(ident="/improve-tutorial/Modeling/Step 9")
```