---
title: "How to..."
format:
  html:
    toc: true
    toc-depth: 2
    toc-title: Contents
    # toc-location: right
    #toc-collapsed: false
    code-fold: false
    number-sections: false
    number-depth: 2
    # smooth-scroll: true
    # message: false
    # warning: false
    # highlight-style: nord
    # output-file: 
    embed-resources: true
    # theme: 
    #   light: [materia, style.scss]
    #   dark: slate
    fontsize: 16px
    # grid:
    # sidebar-width: 300px
    fig-cap-location: top
    fig-align: left
    code-annotations: true
knitr:
    opts_chunk:
      collapse: true
      dev: "ragg_png"
      code-fold: show
# params:
#   include_questions: true
editor_options: 
  markdown: 
    wrap: 72
# project:
#   execute-dir: project
lightbox: true
---


```{r}
#| include: false 
library(improveR)
library(tidyverse, quietly = T)
library(tictoc) #time keeping of functions
options(width=10000) #make results window wide; horizontal scrollbar


fn_connect <- function() {
  Sys.setenv(
    IMPROVER_REPO_URL = "http://envhost1.hc.scintecodev.internal:5310/repository"
    
  )
  Sys.setenv(IMPROVER_AUTO_AUTH = TRUE)

  clearConnectionData()
  improveRtestsupport::improveConnect()

  setEditable(TRUE)
  print(Sys.getenv("IMPROVER_REPO_URL"))
}

fn_connect()

improveConnected()
```


# ... get all steps of an analysis tree

```{r}
#| eval: false
#| code-fold: false

childResourcesATModeling <- improveR::loadChildResources(ident='/improve-tutorial/Modeling/') %>%
  select(data) %>%
  unnest(data) %>%
  select(""name"", ""entityId"", ""path"")

```


# ... get all child steps of a step

Version 1:
```{r}
#| code-fold: false
dfChild <- loadChildSteps(ident="improve-tutorial/Modeling/Step 1")
dfChild$data[[1]] %>%
  select(path, resourceId)
```

Version 2:
```{r}
#| code-fold: false
envStep <- getStep(ident="improve-tutorial/Modeling/Step 1")
envStep$children$load()
ls(envStep$children)
```

# ... get the workflow leading to a step
```{r}
#| code-fold: false
#| eval: false
envATReportingStep1 <- getStep("/improve-tutorial/Reporting/Step 1")

# get the workflow and show details on Step 'Modeling/Step 10'
envATReportingStep1$lineage$load(treeDepth=-1)

#show all steps which form part of the workflow
envATReportingStep1$workflow$df()
```

Note that before calling `lineage$load()`, the `df()` in the `workflow` binding will only return the step itself, i.e.
`Reporting/Step 1`.


# ... get the elements aka resources of a step's inventory

```{r}
#| code-fold: false
#| eval: false

# appraoch 1 
stepInventory <- loadChildResources(ident="/improve-tutorial/Modeling/Step 1")$data[[1]]
ncol(stepInventory)

# approach 2
envStep <- getStep(ident="/improve-tutorial/Modeling/Step 1")
stepInventory2 <- envStep$getStepInventory()$data[[1]]
ncol(stepInventory2)

# the approach with getStep returns two additional columns
setdiff(names(stepInventory2), names(stepInventory)) # 'isVersion', 'inventoryPath'

# approach 3
envStep <- getStep(ident="/improve-tutorial/Modeling/Step 1")
envStep$stepDf #no does not show content of inventory; in the inventory is more than remote files;



```

# ... remove a step from a workflow

```{r}
#| eval: false
#| code-fold: false
#| code-summary: "Remove step from workflow leading a specific step"
step$workflow$removeStep()
```

# ... get the outdated elements of an inventory

```{r}
#| eval: false
#| code-fold: false

envATReportingStep1 <- getStep("/improve-tutorial/Reporting/Step 1")
changedAndOutdatedFiles <- envATReportingStep1$workflow$changedAndOutdatedFiles()

#outdated elements in an inventory
inventory <- envATReportingStep1$getStepInventory()
inventoryOutdatedLinks <- inventory %>%
  select(data) %>%
  unnest(data) %>%
  filter(outdatedLink==TRUE) 
nrow(inventoryOutdatedLinks)

#outdated link in resource
x <- loadResource("/improve-tutorial/Reporting/Step 1")
x$outdatedLink  #outdatedLink can be FALSE even if in the inventory of the Step are outdated elements

```

* elaborate on the difference between outdated elements of an inventory and changedAndOutdated in the workflow context
* outdatedAndChanged is different; also includes changed; changes input is a different one; outdatedLink=>here only outdated links; changed: files are not the same; changed red frame; outdated yellow frame



# ... rerun changed and outdated steps with in a workflow

```{r}
#| eval: false
#| code-fold: false


step$workflow$rerunChangedandOutdated
```


# Common Errors

# Tooling

When defining a step's tooling, no interactive e.g. RStudio instance is permissble. Executing a step via {improveR} requires a tooling
which can be run in batch mode. 

