---
title: "Import & Export"
format:
  html:
    toc: true
    toc-depth: 2
    toc-title: Contents
    toc-location: right
    code-fold: false
    fontsize: 16px
    fig-align: left
knitr:
    opts_chunk:
      collapse: true
      eval: false
---

## Overview

improve supports exporting workflows and folders as portable `.zip` packages that can be imported into other repositories or shared between teams.

## Workflow export/import

### Export a workflow

```{r}
library(improveR)

# Get the workflow from an analysis tree
workflow <- getWorkflow(analysisTree)

# Export to a local folder
exportWorkflow(workflow,
  workflowName = "PK-Analysis-Q1",
  targetFolder = "~/exports")
```

This creates three files:

| File | Content |
|------|---------|
| `PK-Analysis-Q1.zip` | The workflow with all steps, files, and structure |
| `PK-Analysis-Q1LinkMapping.json` | Maps file references between source and target |
| `PK-Analysis-Q1ToolMapping.json` | Maps tool configurations for the target environment |

### Import a workflow

```{r}
# Create a target folder
targetFolder <- createFolder("/Projects", "Imported-Analyses")

# Import the workflow
importWorkflow("~/exports/PK-Analysis-Q1.zip", targetFolder)

# Verify: the analysis tree and all steps are recreated
importedTree <- loadResource("./PK-Analysis-Q1", targetFolder)
importedWorkflow <- getWorkflow(importedTree)
```

The import preserves:

- Step descriptions and rationale
- Step dependencies (parent-child relationships)
- File references and links
- Process configurations

## Folder export/import

For simpler transfers without workflow structure:

### Export a folder

```{r}
exportFolder(sourceFolder,
  exportName = "StudyData",
  targetFolder = "~/exports")
```

### Import a folder

```{r}
targetFolder <- createFolder("/Projects", "ImportedData")
importFolder("~/exports/StudyData.zip", targetFolder)

# Verify contents
children <- loadChildResources(targetFolder)
```

## Incremental import

When re-importing to a location that already has content, the import can handle conflicts:

```{r}
# Import with skip on conflict (keep existing)
importWorkflow("~/exports/PK-Analysis-Q1.zip", targetFolder)
```
