---
title: "Metadata"
format:
  html:
    toc: true
    toc-depth: 2
    toc-title: Contents
    toc-location: right
    code-fold: false
    fontsize: 16px
    fig-align: left
knitr:
    opts_chunk:
      collapse: true
      eval: false
---

## Overview

Metadata in improve attaches structured information to resources — compounds, indications, dates, and other domain-specific attributes. Metadata definitions are configured by administrators; users assign values to resources.

## Metadata definitions

Definitions describe what metadata fields are available and their types.

```{r}
library(improveR)

# Load all available metadata definitions
definitions <- loadMetaDataDefinitions()

# Each definition has a name and type
definitions[, c("name", "metadataType")]
#          name metadataType
# 1    Compound         TEXT
# 2  Indication          LOV
# 3 ProgramStart        DATE
```

Types:

| Type | Description |
|------|-------------|
| `TEXT` | Free text value |
| `LOV` | List of values (pick from predefined options) |
| `DATE` | Date value |

## Adding metadata to a resource

```{r}
# Text value
addMetaDate(folder, "Compound", value = "Aspirin")

# List of values (must match a predefined option)
addMetaDate(folder, "Indication", value = "Pain")

# Date value
addMetaDate(folder, "ProgramStart", value = lubridate::ymd("2026-01-15"))
```

## Reading metadata

```{r}
# Load metadata for a resource
meta <- loadMetaData(folder)

# Get the data
metaData <- meta %>% strip()

# Access specific values
metaData[metaData$descriptorName == "Compound", ]$textValue
# "Aspirin"

metaData[metaData$descriptorName == "Indication", ]$lovText
# "Pain"

metaData[metaData$descriptorName == "ProgramStart", ]$dateValueDate
# "2026-01-15"
```

## Updating metadata

```{r}
updateMetaDate(folder, "Compound", "Ibuprofen")
updateMetaDate(folder, "Indication", "Inflammation")
updateMetaDate(folder, "ProgramStart", lubridate::ymd("2026-03-01"))

# Refresh to see changes
meta <- refreshMetaData(folder)
```

## Deleting metadata

```{r}
deleteMetaDate(folder, "Compound")
```

## Bulk operations

```{r}
# Add multiple metadata values at once
addBulkMetaDate(folder, list(
  list(name = "Compound", value = "Aspirin"),
  list(name = "Indication", value = "Pain")
))
```
