---
title: "Permissions & Access Control"
format:
  html:
    toc: true
    toc-depth: 2
    toc-title: Contents
    toc-location: right
    code-fold: false
    fontsize: 16px
    fig-align: left
knitr:
    opts_chunk:
      collapse: true
      eval: false
---

## Overview

Access control in improve is group-based. You create groups, add users to them, then assign permissions to groups on resources. Permissions can inherit down the folder hierarchy.

## Groups

### Create and manage groups

```{r}
library(improveR)

# Create a group
group <- createGroup("Modelers")

# Add users to the group
allUsers <- users()
addGroupUser(group$id, allUsers$id[allUsers$username == "jane"])
addGroupUser(group$id, allUsers$id[allUsers$username == "bob"])

# Check members
members <- loadGroupUsers(group$id)

# Remove a member
removeGroupUser(group$id, allUsers$id[allUsers$username == "bob"])

# Delete a group
deleteGroup(group$id)
```

## Setting permissions

Permissions are set on resources (folders, files, analysis trees, steps) for groups. Each permission entry specifies:

| Field | Description |
|-------|-------------|
| `visible` | Can see the resource |
| `read` | Can read content |
| `modify` | Can modify content |
| `changeRights` | Can change permissions |
| `inherit` | Propagates to child resources |

### Single permission

```{r}
# Grant read access to a folder
setResourcePermission(folder,
  memberId = group$id,
  visible = TRUE, read = TRUE,
  modify = FALSE, changeRights = FALSE)
```

### Multiple permissions at once

```{r}
# Set multiple permissions (replaces all existing)
replaceResourcePermissions(folder, list(
  list(memberId = adminGroup$id,
       visible = TRUE, read = TRUE, modify = TRUE,
       changeRights = TRUE, inherit = TRUE, rightsArea = 1L),
  list(memberId = readonlyGroup$id,
       visible = TRUE, read = TRUE, modify = FALSE,
       changeRights = FALSE, inherit = TRUE, rightsArea = 1L)
))
```

### Update and remove

```{r}
# Read current permissions
acl <- getResourcePermissions(folder)

# Update a specific entry
updateResourcePermission(folder,
  aclEntryId = acl$id[1],
  modify = TRUE)  # grant modify to first entry

# Remove a specific entry
removeResourcePermission(folder, acl$id[2])
```

## Inheritance

When `inherit = TRUE`, child folders and files receive the same permissions.

```{r}
# Parent folder: full access, inherited
replaceResourcePermissions(parentFolder, list(
  list(memberId = group$id,
       visible = TRUE, read = TRUE, modify = TRUE,
       changeRights = FALSE, inherit = TRUE, rightsArea = 1L)
))

# Child folder automatically gets the same permissions
childPerms <- effectiveUserPermissions(childFolder)
```

When `inherit = FALSE`, the permission applies only to the resource itself:

```{r}
# Read-only at root level, not inherited to children
replaceResourcePermissions(rootFolder, list(
  list(memberId = group$id,
       visible = TRUE, read = TRUE, modify = FALSE,
       changeRights = FALSE, inherit = FALSE, rightsArea = 1L)
))

# Children do NOT get this permission
```

## Checking effective permissions

```{r}
# What permissions does a specific user have?
rights <- effectiveRights(folder, memberId = userId)
rights$read    # TRUE/FALSE
rights$modify  # TRUE/FALSE

# All effective permissions on a resource
perms <- effectiveUserPermissions(folder)
```
