---
title: "Reporting — From Repository to Word and PowerPoint"
format:
  html:
    toc: true
    toc-depth: 2
    toc-title: Contents
    toc-location: right
    code-fold: false
    fontsize: 16px
    fig-align: left
knitr:
    opts_chunk:
      collapse: true
      eval: false
---



## Overview

This article shows how to pull data, figures, and text from the improve
repository and render them into reports. Everything stays traceable —
each output links back to a specific step, version, and audit trail entry.

The examples use the EDA analysis tree at `/demos/DMG-mvp-mcp/EDA` which
contains three finished steps producing summary statistics, covariate
analyses, and outlier reports.

## Connecting



::: {.cell}

```{.r .cell-code}
library(improveR)

# Connect to the repository
improveConnect()
setEditable(TRUE)
```
:::



When running inside a step, `pwd()` returns the current step. Relative
paths resolve from there — use `../` to navigate to other steps or trees.



::: {.cell}

```{.r .cell-code}
# The current step
step <- pwd()
step$path
#> "/demos/DMG-mvp-mcp/Reporting/Step 1"

# Get data from a sibling analysis tree
summary <- getData("../../EDA/Step 1/eda_summary.csv")

# Go up two levels, into another tree's step
plots <- getFile("../../EDA/Step 2/eda_covariate_plots.png")
```
:::



## Loading Files

### loadFile — download to local workspace

`loadFile()` downloads a file from the repository to your working directory
and returns a descriptor. The local file path is in `$data[[1]]`.



::: {.cell}

```{.r .cell-code}
# By relative path — navigate from the current step to another step's output
desc <- loadFile("../../EDA/Step 1/eda_summary.csv")
desc$data[[1]]
#> "/home/user/workspace/eda_summary.csv"

# Read the downloaded file
data <- read.csv(desc$data[[1]])

# By absolute path
desc <- loadFile("/demos/DMG-mvp-mcp/EDA/Step 1/eda_summary.csv")

# By entity ID
desc <- loadFile("envhost1.hc.scintecodev.internal-5310:FI-242425")
```
:::



### addIdToName — handling name collisions

When downloading files from different steps that share the same filename,
the entity ID is prepended by default to keep them distinct.



::: {.cell}

```{.r .cell-code}
# Default: entity ID prepended
desc <- loadFile("../../EDA/Step 1/eda_summary.csv", addIdToName = TRUE)
basename(desc$data[[1]])
#> "envhost1.hc.scintecodev.internal-5310_FI-242425_eda_summary.csv"

# Plain filename
desc <- loadFile("../../EDA/Step 1/eda_summary.csv", addIdToName = FALSE)
basename(desc$data[[1]])
#> "eda_summary.csv"
```
:::



### Loading a specific version

Every file in improve has a version history. The entity version ID is the
entity ID followed by the version number: `entityId-version`.



::: {.cell}

```{.r .cell-code}
# Check the current version
res <- loadResource("../../EDA/Step 1/eda_summary.csv")
res$entityId
#> "envhost1.hc.scintecodev.internal-5310:FI-242425"
res$entityVersionId
#> "envhost1.hc.scintecodev.internal-5310:FI-242425-1"

# Load a specific version
desc <- loadFile("envhost1.hc.scintecodev.internal-5310:FI-242425-1")
```
:::



### Links vs copies

When a step references a file, it can be a **link** or a **copy**:

- **Link** (`asLink = TRUE`, default): the step references the original.
  The step always sees the current version of the source. The link itself
  cannot be edited — only the source can be updated.
- **Copy** (`asLink = FALSE`): the step has its own independent version.
  Changes to the original do not propagate, and the copy can be updated
  independently.



::: {.cell}

```{.r .cell-code}
template <- createStepTemplateEnv(treeIdent = myTree)

# Command file as a copy — the step owns this file
template$addStepRemoteFile(
  "./scripts/analysis.R",
  variableName = "command-file",
  asLink = FALSE
)

# Input data as a link — always uses the current source version
template$addStepRemoteFile(
  "./data/raw_data.csv",
  name = "input_data.csv",
  asLink = TRUE
)
```
:::



## Typed File Access

### getData — parsed tables

`getData()` downloads a file, auto-detects the format, and parses it. The
parsed data frame is in `$data[[1]]`.

Supported formats: `.csv`, `.xlsx`/`.xls`, `.rds`, `.sas7bdat`. A custom
parser can be provided via the `parser` argument.



::: {.cell}

```{.r .cell-code}
# CSV — auto-detected, pulling from another step
desc <- getData("../../EDA/Step 1/eda_summary.csv")
desc$dataType
#> "CSV"

params <- desc$data[[1]]
head(params)
#>   DOSE N_SUBJECTS N_OBS DV_MEAN DV_MEDIAN DV_MIN DV_MAX CV_PCT
#> 1   25          6    62    5.44       5.5      0   11.5   58.0
#> 2   50          6    65   10.84      11.9      0   20.8   58.3

# Custom parser
desc <- getData("./custom_format.dat",
                parser = function(path) read.table(path, header = TRUE, sep = "\t"))
```
:::



### getFile — file descriptor with metadata

`getFile()` downloads a file and returns a descriptor with metadata. The
local file path is in `$path`.



::: {.cell}

```{.r .cell-code}
desc <- getFile("../../EDA/Step 1/eda_summary.csv")
desc$path
#> "/home/user/workspace/eda_summary.csv"
desc$caption
#> "Source (Entity version ID):...-5310:FI-242425-1\nLast modified on: 2026-04-24 12:10:54"
desc$entityId
#> "envhost1.hc.scintecodev.internal-5310:FI-242425"
```
:::



### getCopy — plain local copy

`getCopy()` downloads the file without prepending the entity ID and without
creating a link in the step inventory.



::: {.cell}

```{.r .cell-code}
desc <- getCopy("../../EDA/Step 1/eda_summary.csv")
basename(desc$path)
#> "eda_summary.csv"
```
:::



### getTextString — text as character string

`getTextString()` reads a text file and returns the content as a single
character string in `$data`.



::: {.cell}

```{.r .cell-code}
desc <- getTextString("../../EDA/Step 1/eda_summary.R")
cat(desc$data)
#> # EDA — Summary Statistics Table
#> # Input:  data.csv (linked from Data Assembly Step 1 output)
#> ...
```
:::



### includeGraphics — figures in Rmd/Quarto

`includeGraphics()` downloads an image and returns a `knitr` include
directive. Use it directly inside an Rmd or Quarto code chunk.



::: {.cell}

```{.r .cell-code}
# In an Rmd chunk:
includeGraphics("../../EDA/Step 2/eda_covariate_plots.png",
                caption = "Covariate distributions by dose group")
```
:::



### getFilesFromFolder — listing files

`getFilesFromFolder()` lists all files (and links) in a folder. Use
`filePattern` to filter by name and `recurse = TRUE` to include subfolders.



::: {.cell}

```{.r .cell-code}
# All files in current step
files <- getFilesFromFolder(".")
files$name

# Filter by pattern
csvFiles <- getFilesFromFolder(".", filePattern = "\\.csv$")

# Recurse into subfolders
allFiles <- getFilesFromFolder("/demos/DMG-mvp-mcp/EDA", recurse = TRUE)
```
:::



## Rendering a Word Report

An Rmd file that pulls EDA outputs from the repository and renders to Word:

````markdown
---
title: "EDA Report"
output: word_document
---

```{r setup, include=FALSE}
library(improveR)
improveConnect()
```

## Summary Statistics

```{r}
summary <- getData("/demos/DMG-mvp-mcp/EDA/Step 1/eda_summary.csv")
knitr::kable(summary$data[[1]], caption = "PK Summary by Dose Group")
```

## Covariate Analysis

```{r}
includeGraphics("/demos/DMG-mvp-mcp/EDA/Step 2/eda_covariate_plots.png",
                caption = "Covariate Distributions")
```

## Outlier Analysis

```{r}
outliers <- getData("/demos/DMG-mvp-mcp/EDA/Step 3/eda_outliers.csv")
knitr::kable(outliers$data[[1]], caption = "Identified Outliers")
```
````

Render:



::: {.cell}

```{.r .cell-code}
rmarkdown::render("eda_report.Rmd", output_format = "word_document")
```
:::



### Using a Word template from the repository



::: {.cell}

```{.r .cell-code}
getCopy("/demos/reportArticle/template.docx")

rmarkdown::render("eda_report.Rmd",
                  output_format = rmarkdown::word_document(
                    reference_doc = "template.docx"
                  ))
```
:::



### Rendering to PowerPoint



::: {.cell}

```{.r .cell-code}
rmarkdown::render("eda_slides.Rmd",
                  output_format = rmarkdown::powerpoint_presentation())
```
:::



## Running a Report Step

Reports are typically generated inside improve steps. The step pulls
data from prior analysis steps, renders the report, and checks it back
into the repository as a versioned output.



::: {.cell}

```{.r .cell-code}
# Run the report step with a clean inventory
runStepResource("/demos/DMG-mvp-mcp/EDA/ReportStep", resetInventory = TRUE)

# Wait for completion
finishRunResource("/demos/DMG-mvp-mcp/EDA/ReportStep")

# The rendered report is now in the step inventory
inv <- loadFullChildResources("/demos/DMG-mvp-mcp/EDA/ReportStep")$data[[1]]
reports <- inv[grepl("\\.docx$", inv$name), ]
reports$name
```
:::



## Shiny App with improve Login

For interactive visualisation, a Shiny app can authenticate against improve
using OAuth device code flow. A complete working example is available at
`inst/examples/shiny-login-app.R`.

After successful authentication, all improveR functions work within the
Shiny session:



::: {.cell}

```{.r .cell-code}
library(shiny)
library(improveR)

server <- function(input, output, session) {
  # After OAuth login completes:
  # improveConnect() is called with the obtained token

  output$summary_table <- renderTable({
    req(improveConnected())
    desc <- getData("/demos/DMG-mvp-mcp/EDA/Step 1/eda_summary.csv")
    desc$data[[1]]
  })

  output$covariate_plot <- renderImage({
    req(improveConnected())
    desc <- getFile("/demos/DMG-mvp-mcp/EDA/Step 2/eda_covariate_plots.png")
    list(src = desc$path, contentType = "image/png")
  }, deleteFile = FALSE)
}
```
:::

