---
title: "Reviews"
format:
  html:
    toc: true
    toc-depth: 2
    toc-title: Contents
    toc-location: right
    code-fold: false
    fontsize: 16px
    fig-align: left
knitr:
    opts_chunk:
      collapse: true
      eval: false
---

## Overview

Reviews in improve follow a structured workflow where a **requestor** creates a review, assigns **reviewers**, and submits resources for inspection. Reviewers accept the invitation, examine individual entries, and approve or reject them.

### Review statuses

| Status | Description |
|--------|-------------|
| **Planning** | Initial state — setting up entries and reviewers |
| **Reviewing** | Active review — reviewers can comment, approve, reject |
| **Approved** | Locked — review and all reviewed resources are locked |

### Transitions

```
Planning  →  Reviewing  →  Approved
   ↑             ↓
   └── revert ───┘
```

## Complete review flow

### 1. Create a review

The requestor creates a review with files to be reviewed and assigns reviewers.

```{r}
library(improveR)

# Create a review with a file and a reviewer
review <- createReview(
  name = "Data Analysis Review Q1",
  parentIdent = "/Projects/Reviews",
  comment = "Review of Q1 analysis outputs",
  templateId = NULL,
  resourceIds = list(analysisFile$resourceId),
  reviewerIds = list(reviewerUserId),
  dueDate = "2026-06-30"
)
```

You can add more entries and reviewers after creation:

```{r}
# Add another file to the review
createReviewEntry(review, resourceIds = list(additionalFile$resourceId))

# Add another reviewer
createReviewer(review, userId = secondReviewerId, username = "jane")
```

### 2. Reviewer accepts the invitation

The reviewer must accept the invitation while the review is in **Planning** state.

```{r}
# As the reviewer:
acceptReviewInvitation(review, comment = "Happy to review")
```

### 3. Start the review

The requestor transitions the review to **Reviewing**.

```{r}
# As the requestor/admin:
changeReviewStatus(review, "Reviewing")
```

### 4. Reviewer works on entries

The reviewer can now comment on entries and approve or reject them individually.

```{r}
# As the reviewer:

# Add a comment to a specific file
createReviewComment(
  ident = review,
  resourceIdent = analysisFile$resourceId,
  comment = "The statistical model looks correct",
  commentType = "GENERAL"
)

# Get all entries
entries <- getReviewEntries(review)

# Approve one entry
approveReviewEntries(review,
  reviewEntryIds = list(entries$id[1]),
  comment = "Verified and approved"
)

# Reject another entry
rejectReviewEntries(review,
  reviewEntryIds = list(entries$id[2]),
  comment = "Missing validation step, please revise"
)

# Reset a decision (e.g. after re-review)
resetReviewEntries(review,
  reviewEntryIds = list(entries$id[2]),
  comment = "Re-reviewing after revision"
)
```

### 5. Remove entries

Individual entries can be removed by their entry IDs:

```{r}
entries <- getReviewEntries(review)
deleteReviewEntry(review, reviewEntryIds = list(entries$id[2]))
```

### 6. Finalize the review

Once all entries are approved (no rejections), the requestor can lock the review:

```{r}
# As the requestor/admin:
changeReviewStatus(review, "Approved")
```

This locks both the review and all reviewed resources. Only an admin can revert to Reviewing.

## Reading review data

```{r}
# Load a review by ID
review <- getReviewById(reviewIdent)

# List all reviews
allReviews <- loadReviews()

# Get reviewers, entries, comments
reviewers <- getReviewers(review)
entries <- getReviewEntries(review)
comments <- getReviewComments(review)
```
